Efficient Coalition Formation for Web Services

### 2013-03-17 papers/coalition/coalition++.tex
Changes

In this scenario, we assume the existence of a typical community
managed by its master web service and web services need to join it
to be able to get requests from the master. The community master
is characterized by a requests rate $(R_{M})$ from users. Each web
service comes with a given QoS ($QoS_{ws}$) from which the
throughput $T_{ws}$ is excluded. This exclusion allows us to build
our analysis on the particular value of $T_{ws}$. Throughput is
the average rate of tasks a web service can perform per time unit.
Therefore, the master web service is providing tasks with the rate
of $R_{ws}$ and web services perform tasks with average output
quality of $QoS_{ws}$ and a throughput rate of $T_{ws}$. In this
setting, we define the value of the coalition of the web services
within a community as a function increasing in both $T_{ws}$ and
$QoS_{ws}$ as follows:
\begin{equation*}
output(C) = \sum_{ws \in C}{(T_{ws} \times QoS_{ws})}
\end{equation*}
\setlength{\abovedisplayshortskip}{2pt}
\begin{equation}
v(C) = \left\{
  \begin{array}{l l}
    output(C) & \quad \text{if $\sum_{ws}{T_{ws}} \leq R_{M}$}\\
    output(C) \times \frac{R_{M}}{\sum_{ws}{T_{ws}}} & \quad \text{if $\sum_{ws}{T_{ws}} > R_{M}$}
  \end{array} \right.
\end{equation}

@@

The membership decision is made based on throughput and \emph{QoS}
of the considered web service. The goal is to have quality web
services in the community so it stays stable and no other web
services would have incentives to deviate and leave the coalition
$C$. Therefore, a basic method would be to check the core of the
coalition $C$ considering all the current community members (all
web services already residing within the community) and the new
web service. This algorithm uses the \emph{Shapley value}
distribution method as described in Equation \ref{eq:shapley} to
distribute the gain of $v(C)$ among all the members and then
checks if the \emph{Shapley value} payoff vector for this
community having the characteristic function $v(C)$ is in the
\emph{core}. In the \emph{Shapley value} payoff vector, the payoff
for each web service $ws_i$ is calculated based on its marginal
contribution $v(C \cup {i}) - v(C)$ over all the possible
different permutations in which the coalition can be formed, which
makes the payoff distribution fair. Because of going through all
the possible permutations of subsets of $N$, the nature of the
\emph{Shapley value} is combinatorial. This makes it impractical
to use as the size of our coalitions grows. However, it is proven
that in convex games, the \emph{Shapley value} lies in the core
\cite{DBLP:conf/ijcai/GrecoMPS11, myerson1991game}. Thus, if the
\emph{Core} is non-empty, the payoff vector is a member of the
\emph{Core}. The following proposition is important to make our
algorithm tractable.
% so in our algorithm we check the core membership of this payoff vector.

@@

As a direct consequence of this proposition, in order to keep the
characteristic function convex, new web services should have more
marginal contribution as the coalition size grows.

@@

\textbf{Jamal: I stopped Here}

@@

In this scenario, we do not consider pre-defined communities and master web services. Here web services are interested in forming coalitions to perform better and share benefits. We do not consider concept of master web services providing requests for others, each web service comes with its own request load which in this scenario we call it \emph{market share}. Therefore coalitions will form is web services working together can perform and gain more than working alone individually. The worth or value of coalitions of web services is different in this setting. We average the $quality$ for all metrics of web service $(Q^{m})$, multiply it by a weight where the importance or weigth of each metric is dependent on type of users and requests web services are dealing with. For each web service we multiply this by web service's market share $(MS_{ws})$ and the summation would be the characteristic or valuation of independent web service coalitions.

\begin{equation}
v(C) = \sum_{ws \in C}{ \Big( \frac{\sum_{m=metrics}{Q^m_{ws} \times w_{m}}}{|metrics|} \Big) \times MS_{ws} }
\end{equation}

%$CC^{m}$ is \emph{Collabrative Coefficient} for each QoS metric (non-functional parameter), it determines how web services providing that QoS metric individually will be effected in a coalition environment. If it is more than 1, is means they increase performance by working together and they are \emph{super additive}, otherwise it depicts they would perform better in linear sense if they worked alone.
$w_m$ is the weight of important for a \emph{QoS} metric, it gives the community the flexibility to emphasize or neglect a specific \emph{QoS} parameter. Note that $ \sum_{m}{w_m} = 1 $. In this scenario we have no master web services, we assume they either dynamically vote for one, or collaborate and agree on decisions and distribute task in round robin or fair fashion. Here, we utilized the same merge-and-split algorithm which we applied for the \emph{``Web Services and Many Communities''} scenario, to find stable coalitions of web serivces.

@@

\section{Experimental Results}\label{s:resutls}

In this section, we discuss the experiments we performed for our scenarios to validate applicability and performance of our proposed methods in realistic environments. We created a pool of web services and populated most of their \emph{QoS} parameters from a real world web service dataset \cite{DBLP:conf/smc/Al-MasriM09a}. We programmed the simulations using Java and executed the simulations on a Intel Xeon X3450 machine with 4GBs of memory. For other parameters missing from the dataset, we used normal random distribution with parameters that make most sense in real world examples.

We analyze the most important performance result the communities should perform, quality and quantity of tasks successfully performed by the communities. This is the most important criteria for user satisfaction. We initiate the community with few web services, then let rejecting and accepting of random web services go for a short number of iterations, after that we start task distribution for the communities and let them allocate tasks for web services. Then we measure the average output performance of tasks in communities following different methods.

\begin{figure}[!t]
\centering
%\includegraphics[scale=0.6]{graph1Final+.eps}
\includegraphics[width=3in]{task_done_opt.eps}
\includegraphics[width=3in]{task_qos_opt.eps}
\caption{Part (a): Cumulative number of tasks succesfully done. Part
(b): Average QoS of tasks performed.} \label{performanceall}
\end{figure}

Figure \ref{performanceall} depicts the results of optimal \emph{$\epsilon$-core}, by assigning $\epsilon$ to 15\% of total community worth, $\epsilon = 0.15 \times v(C)$, allowing subsets of coalition to gain maximum 15\% of $v(C)$, \emph{Depth-1 Convex-Checker}, \emph{Depth-2 Convex-Checker}, \emph{3-Way Satisfactory}\cite{DBLP:conf/IEEEscc/LimTMB12} and \emph{2 Player Non-Cooperative}\cite{DBLP:conf/IEEEscc/KhosravifarABT11} methods in \emph{one grand community with many web services} scenario. In the \emph{optimal $\epsilon$-core} method we capped the coalition size to 25 web services, since anything more than that would require analysis with time complexity of $O(2^{n-1})$, making it impractical to run in our simulations. In other methods there were no cap on size of the community and we had communities of size 60 web services at some points. The results show \emph{depth-2 and 1 convex checker} methods are performing better compared to other methods and their performance is close to optimal \emph{$\epsilon$-core} method.

@@

As mentioned in section \ref{s:preliminaries}, the concept of \emph{core}, assumes no coalition of players can gain anything by deviating which is a fairly strong requirement, that is why the notion of \emph{$\epsilon$-core} was introduced. Least-Core $e(G)$ of a game $G$, is the minimum amount of $\epsilon$ in which the core is not empty. We evaluated the least-core value using our valuation function and our set of web services. We pick random number of web services from our dataset and formed around 10,000 random coalitions consisting of 3 to 26 web services. We choose 26 as maximum number of members in our grand coalition since it is computationally very complex for bigger coalitions to compute \emph{$\epsilon$-core}. We set $\epsilon$ to a ratio of total payoff of coalition $v(C)$ between 0 and 1. The results in figure \ref{f_leastcore} illustrates, almost 10\% of our random web service coalitions have non-empty \emph{core} solution and \emph{$\epsilon$-core} solution is \emph{always} non-empty when we let agents gain 30\% (of $v(C)$) more, by deviating.

One of the properties of coalition structure formation algorithms in our scenario is that they partition web services with low throughput rate usually join coalitions with less request rate. Since the characteristic function $v(C)$ and  the fair payoff vector of Shapley value is proportional to web services' contribution, the web services with small contribution will get paid much less in communities having web services with high throughput. On the other hand, according to valuation function $v(C)$ web services with high throughput will not contribute well to communities with low amount of user requests. The strong web services are much more likely to deviate weak coalitions, joining a stronger coalition, making the initial coalition unstable.

@@

In figure \ref{performancemany} we compare our \emph{Web Services and many Communities} scenario with a method which ignores QoS parameters and forms coalitions and allow web services to join only if they have enough requests for them in other words web services can join a community when request rate is less that throughput of all web services inside community ignoring QoS parameters. We name this method \emph{Random Formation} and use it as our benchmark for our QoS aware coalition formation process. As results illustrate clearly, our QoS aware method, forms better coalitions of web services improving performance and satisfaction for both web service and coalitions. 
% An example of a floating figure using the graphicx package.
% Note that \label must occur AFTER (or within) \caption.
% For figures, \caption should occur after the \includegraphics.
% Note that IEEEtran v1.7 and later has special internal code that
% is designed to preserve the operation of \label within \caption
% even when the captionsoff option is in effect. However, because
% of issues like this, it may be the safest practice to put all your
% \label just after \caption rather than within \caption{}.
%
% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
% option should be used if it is desired that the figures are to be
% displayed while in draft mode.
%
%\begin{figure}[!t]
%\centering
%\includegraphics[width=2.5in]{myfigure}
% where an .eps filename suffix will be assumed under latex,
% and a .pdf suffix will be assumed for pdflatex; or what has been declared
% via \DeclareGraphicsExtensions.
%\caption{Simulation Results}
%\label{fig_sim}
%\end{figure}

@@

In this paper, we proposed a cooperative game theory based model for aggregation of web services within communities. The goal of our services is to maximize efficiency of performing tasks by collaborating and forming stable communities. Our method considers stability and fairness for all web services within a community and offers an applicable mechanism for membership requests and selection of web services. The goal is to increase revenue by improving user satisfaction, which comes from the ability to perform tasks with higher quality and quantity. Simulation results show that our, polynomial in complexity, approximation algorithms provide web services and community owners with applicable and near-optimal decision making mechanisms.

As future work, we would like to perform more analytical and theoretical work on the convexity condition for the characteristic function on web service applications. It is possible to work on compact or succinct representation of our service model (i.e. a graph representation) to reduce the input complexity and solution concepts of cooperative game and core membership checking algorithms. From web service aspect, the work can be extended to consider web service compositions where a group of web services having different set of skills cooperate to perform composite tasks. Also bargaining theory from cooperating game theory concepts can be used in our scenario. The web services would stay and join communities as they wish, however, they would resolve the instability and unfairness issues by side payments.

>>>>

In this scenario, we assume the existence of a typical community
managed by its master, and web services need to join it to be able
to get requests from the master. The community master is
characterized by a requests rate $(R_{M})$ from users. Each web
service comes with a given QoS ($QoS_{ws}$) from which the
throughput $T_{ws}$ is excluded. This exclusion allows us to build
our analysis on the particular value of $T_{ws}$. Throughput is
the average rate of tasks a web service can perform per time unit.
Thus, web services perform tasks with an average output quality of
$QoS_{ws}$ and a throughput rate of $T_{ws}$. In this setting, we
define the value of the coalition of the web services within a
community as a function increasing in both $T_{ws}$ and $QoS_{ws}$
as follows:
\begin{equation*}
output(C) = \sum_{ws \in C}{(T_{ws} \times QoS_{ws})}
\end{equation*}
\setlength{\abovedisplayshortskip}{2pt}
\begin{equation}
v(C) = \left\{
  \begin{array}{l l}
    output(C) & \quad \text{if $\sum_{ws}{T_{ws}} \leq R_{M}$}\\
    output(C) \times \frac{R_{M}}{\sum_{ws}{T_{ws}}} & \quad \text{if $\sum_{ws}{T_{ws}} > R_{M}$}
  \end{array} \right.
\end{equation}

@@

The membership decision is made based on throughput and \emph{QoS}
of the considered web service. The goal is to have quality web
services in the community so it stays stable and no other web
services would have incentives to deviate and leave the coalition
$C$. Therefore, a basic method would be to check the core of the
coalition $C$ considering all the current community members (all
web services already residing within the community) and the new
web service. This algorithm uses the \emph{Shapley value}
distribution method as described in Equation \ref{eq:shapley} to
distribute the gain of $v(C)$ among all the members and then
checks if the \emph{Shapley value} payoff vector for this
community having the characteristic function $v(C)$ is in the
\emph{core}. In the \emph{Shapley value} payoff vector, the payoff
for each web service $ws_i$ is calculated based on its marginal
contribution $v(C \cup {i}) - v(C)$ over all the possible
different permutations in which the coalition can be formed, which
makes the payoff distribution fair. Because of going through all
the possible permutations of subsets of $N$, the nature of the
\emph{Shapley value} is combinatorial, which makes it impractical
to use as the size of our coalitions grows. However, it is proven
that in convex games, the \emph{Shapley value} lies in the core
\cite{DBLP:conf/ijcai/GrecoMPS11, myerson1991game}. Thus, if the
\emph{Core} is non-empty, the payoff vector is a member of the
\emph{Core}. The following proposition is important to make our
algorithm tractable.
% so in our algorithm we check the core membership of this payoff vector.

@@

Thus, in order to keep the characteristic function convex, new web
services should have more marginal contribution as the coalition
size grows.

@@

@@

In this scenario, there is no pre-defined community and web
services are interested in forming new ones to perform better and
share benefits (see Figure \ref{fig_sim3}). Consequently, there is
no master web service providing requests for others, but each web
service comes with its own request load called \emph{market
share}. Communities can be formed because by working together web
services can expect to gain more than working individually. The
worth or value of a coalition of web services is different in this
setting. Let $metrics$ be the set of quality metrics, we average
the values $Q^m_{ws}$ for all quality metrics ($m \in metrics$)
for each web service $ws$, multiply it by a weight reflecting the
importance of each metric for each web service ($w^m_{ws}$) (Note
that $ \sum_{m}{w^m_{ws}} = 1 $). For each web service, we
multiply this weighted average by web service's market share
$(\mu_{ws})$ and the summation over all the web services of the
community to be established would be the characteristic or
valuation function. This function has the property to be
increasing in the market share and QoS.

\begin{equation}
v(C) = \sum_{ws \in C}{ \Big( \frac{\sum_{m \in metrics}{Q^m_{ws}
\times w^m_{ws}}}{|metrics|} \Big) \times \mu_{ws} }
\end{equation}

%$CC^{m}$ is \emph{Collabrative Coefficient} for each QoS metric (non-functional parameter), it determines how web services providing that QoS metric individually will be effected in a coalition environment. If it is more than 1, is means they increase performance by working together and they are \emph{super additive}, otherwise it depicts they would perform better in linear sense if they worked alone.
As there is no pre-defined master web services in this scenario,
we assume that the members of each coalition either dynamically
vote for one, or collaborate and distribute tasks in round robin
or fair fashion. To form stable coalitions, we use the same
merge-and-split algorithm we applied for the \emph{``Web Services
and Many Communities''} scenario.

@@

\section{Experimental Results and Analysis}\label{s:resutls}

In this section, we discuss the experiments we performed for our
scenarios to validate the applicability and performance of our
proposed methods in realistic environments. We created a pool of
web services and populated most of their \emph{QoS} parameters
from a real world web service dataset
\cite{DBLP:conf/smc/Al-MasriM09a}. We implemented the simulations
using Java and executed the simulations on an Intel Xeon X3450
machine with 4GBs of memory. For other parameters missing from the
dataset, we used normal random distribution with parameters
estimated using the method of maximum likelihood.

One of the key criteria reflecting the performance of web service
coalitions is the user satisfaction. User satisfaction can be
measured in terms of quality and quantity of requests (or tasks)
successfully answered by the communities. We initiate the
communities with few web services, then let rejecting and
accepting random web services go for a short number of iterations.
After that, we start the request distribution for the communities
and let them allocate requests among member web services.
Thereafter, we measure the average output performance of tasks in
communities following different methods.

\begin{figure}[!t]
\centering
%\includegraphics[scale=0.6]{graph1Final+.eps}
\includegraphics[width=3in]{task_done_opt.eps}
\includegraphics[width=3in]{task_qos_opt.eps}
\caption{Part (a): Cumulative number of requests successfully
done. Part (b): Average QoS of requests performed.}
\label{performanceall}
\end{figure}

Figure \ref{performanceall} depicts the results of optimal
\emph{$\epsilon$-core}, \emph{Depth-1 Convex-Checker},
\emph{Depth-2 Convex-Checker}, \emph{3-Way Satisfaction}
\cite{DBLP:conf/IEEEscc/LimTMB12}, and \emph{2 Player
Non-Cooperative} \cite{DBLP:conf/IEEEscc/KhosravifarABT11} methods
in \emph{one grand community with many web services} scenario. For
\emph{$\epsilon$-core}, we assign $\epsilon$ to 15\% of total
community worth, $\epsilon = 0.15 \times v(C)$, which allows
subsets of the coalition to gain maximum 15\% of $v(C)$. In the
\emph{optimal $\epsilon$-core} method, we capped the coalition
size to 25 web services, since anything more than that would
require analysis with time complexity of $O(2^{n-1})$, making it
impractical to run in our simulations. In the other methods, there
were no cap on size of the community and we had communities of
size 60 web services at some points. The results show that our
\emph{depth-2 convex checker} method is performing better compared
to the other methods and its performance is close to optimal
\emph{$\epsilon$-core} method. Our \emph{depth-1 convex checker}
and \emph{3-Way Satisfaction} are also performing well.

@@

As mentioned in Section \ref{s:preliminaries}, the concept of
\emph{core}, assumes no coalition of players can gain anything by
deviating, which is a fairly strong requirement, and that is why
the notion of \emph{$\epsilon$-core} was introduced. Least-Core
$e(G)$ of a game $G$, is the minimum amount of $\epsilon$ so that
the core is not empty. We evaluated the least-core value using our
valuation function and our set of web services. We picked random
number of web services from the dataset and formed around 10,000
random coalitions consisting of 3 to 26 web services. We choose 26
as the maximum number of members in our grand coalition since it
is computationally very complex for bigger coalitions to compute
\emph{$\epsilon$-core}. We set $\epsilon$ to a ratio of total
payoff of coalition $v(C)$ between 0 and 1. The results in Figure
\ref{f_leastcore} illustrates that almost 10\% of our random web
service coalitions have non-empty \emph{core} solution and
\emph{$\epsilon$-core} solution is \emph{always} non-empty when we
let agents gain 30\% more of $v(C)$ by deviating.

One of the properties of coalition structure formation algorithms
in our third scenario is that they partition web services with low
throughput rate so that they usually join coalitions with less
request rate. Since the characteristic function $v(C)$ and the
fair Shapely payoff vector is proportional to web services'
contribution, the web services with small contribution will get
paid much less in communities having web services with high
throughput. On the other hand, according to the valuation function
$v(C)$, web services with high throughput will not contribute well
to communities with low amount of user requests (low market
share). The strong web services are likely to deviate from weak
coalitions, joining a stronger one, which makes the initial
coalition unstable.

@@

In Figure \ref{performancemany}, we compare our \emph{Web Services
and many Communities} scenario with a method which ignores QoS
parameters and forms coalitions by allowing web services to join
only if they have enough requests for themselves. In other words,
web services can join a community when the request rate is less
than the throughput of all the member web services. We name this
method \emph{Random Formation} and use it as a benchmark for our
QoS-aware coalition formation process. As the results illustrate,
our method forms better coalitions of web services improving
performance and satisfaction for both web services and coalitions.
% An example of a floating figure using the graphicx package.
% Note that \label must occur AFTER (or within) \caption.
% For figures, \caption should occur after the \includegraphics.
% Note that IEEEtran v1.7 and later has special internal code that
% is designed to preserve the operation of \label within \caption
% even when the captionsoff option is in effect. However, because
% of issues like this, it may be the safest practice to put all your
% \label just after \caption rather than within \caption{}.
%
% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
% option should be used if it is desired that the figures are to be
% displayed while in draft mode.
%
%\begin{figure}[!t]
%\centering
%\includegraphics[width=2.5in]{myfigure}
% where an .eps filename suffix will be assumed under latex,
% and a .pdf suffix will be assumed for pdflatex; or what has been declared
% via \DeclareGraphicsExtensions.
%\caption{Simulation Results}
%\label{fig_sim}
%\end{figure}

@@

In this paper, we proposed a cooperative game theory-based model
for the aggregation of web services within communities. The goal
of our services is to maximize efficiency by collaborating and
forming stable coalitions. Our method considers stability and
fairness for all web services within a community and offers an
applicable mechanism for membership requests and selection of web
services. The ultimate goal is to increase revenue by improving
user satisfaction, which comes from the ability to perform more
tasks with high quality. Simulation results show that our,
polynomial in complexity, approximation algorithms provide web
services and community owners with applicable and near-optimal
decision making mechanisms.

As future work, we would like to perform more analytical and
theoretical analysis on the convexity condition for the
characteristic function on web service applications. It is
possible to work on compact or succinct representation of our
service model (i.e. a graph representation) to reduce the input
complexity and solution concepts of cooperative game and core
membership checking algorithms. From web service perspective, the
work can be extended to consider web service compositions where a
group of web services having different set of skills cooperate to
perform composite tasks. Also bargaining theory from cooperating
game theory concepts can be used to help web services resolve the
instability and unfairness issues by side payments.
